---
title: "Clean Up and Test Democracy Mentions Data"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
bibliography: source_refs.bib
output: 
  html_document: 
    keep_md: yes
    self_contained: no
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      cache.rebuild = FALSE,
                      message = FALSE,
                      warning = FALSE)

```

## Raw data description

The raw data consists of all the mentions of democracy in constitutional documents I was able to extract from the texts gathered by the Comparative Constitutions Project as of 2015. 

### Loading and cleanup

```{r loading}
library(tidyverse)
library(stringr)
library(knitr)

democracy_mentions <- read_csv("../../../../Data/Democracy mentions/democracy.mentions.csv",
                               col_types = cols(YEAR = col_number(),
                                                earliest.dem = col_number(),
                                                first.constitution.in.force = col_number())) %>%
  select(COUNTRY, YEAR, Name, EVENTTYPE, mentions) 

# Some years have a decimal, which we eliminate
democracy_mentions$YEAR <- as.integer(democracy_mentions$YEAR) 

# This is what the mentions look like

democracy_mentions %>%
  select(COUNTRY,YEAR,Name,mentions) %>%
  filter(!is.na(YEAR)) %>%
  head() %>%
  kable()

```

Once the data is loaded, we need to standardize the country names and make sure they correspond to the right country-names in the standard "international system":

```{r convert_to_gw}

library(lubridate)

countries <- democracy_mentions %>% 
  distinct(COUNTRY, YEAR) %>%
  PoliticalDatasets::to_gw_system(country_col = "COUNTRY",
                                  date_col = "YEAR",
                                  match_condition = c("!(COUNTRY == 'Vietnam (Annam/Cochin China/Tonkin)' & country_name =='China' )",
                                              "!(COUNTRY == 'Yugoslavia (Serbia)' & country_name =='Yugoslavia' & (YEAR < 1900 | YEAR == 2006))",
                                              "(match_score == max(match_score))"))

```

The country names for some documents not matched, simply because the documents do not refer to independent countries, or the document does not contain a proper country name: 

```{r not_matched}

anti_join(democracy_mentions,countries) %>% 
  group_by(COUNTRY, Name) %>% 
  summarise(min(YEAR),max(YEAR),n()) %>%
  knitr::kable()

```

We also need to clean up column names, trim whitespace, and eliminate duplicates:

```{r further_cleanup}

democracy_mentions <- countries %>%
  left_join(democracy_mentions) %>% 
  rename(year = YEAR, constitutional.event = EVENTTYPE, ccp_country = COUNTRY) %>% 
  mutate(Name = str_trim(Name), 
         Name = str_replace(Name,"\\.[0-9]$",""),
         constitutional.event = factor(constitutional.event)) %>%
  select(country_name, 
         GWn, 
         year,
         Name,
         constitutional.event,
         mentions,
         ccp_country,
         GWc:in_system) 

# We eliminate duplicates

nrow(democracy_mentions)

democracy_mentions <- distinct(democracy_mentions)

nrow(democracy_mentions)

```

These countries had their names changed:

```{r names_changed}

democracy_mentions %>% 
  filter(country_name != ccp_country) %>% 
  group_by(country_name, ccp_country) %>%
  summarise(min(year), max(year), n()) %>%
  knitr::kable()
  
```

And here's what the data looks like, per country:

```{r country_plot, fig.height=25}

library(forcats)

ggplot(data = democracy_mentions) +
  geom_count(aes(x=fct_rev(reorder(country_name,year,FUN = min)), y = year, color = fct_infreq(constitutional.event))) +
  coord_flip() +
  labs(x = "", color = "Constitutional\nevent", size = "Number of \ntext snippets") +
  theme_bw() 

```

These are the countries where the first mention is not "new" or "interim" constitution:

```{r not_new}

democracy_mentions %>% 
  group_by(country_name) %>%
  filter(!(first(constitutional.event) %in% c("new","interim"))) %>% 
  summarise(year = first(year), Name = first(Name), constitutional.event = first(constitutional.event)) %>%
  knitr::kable()

```

These countries need to be fixed:

```{r fixing_countries}

fixed_countries <- democracy_mentions %>% 
  group_by(country_name) %>%
  filter(!(first(constitutional.event) %in% c("new","interim"))) %>% 
  summarise_all(funs(first(.))) %>%
  mutate(year = c(1959, 1959, 1908, 1981, 1948, 1957, 
                  1968, 1978, 1959, 1875, 1962, 1970),
         constitutional.event = c("new","new","new","new","new","new",
                                  "new","new","new","new","new","new"))

democracy_mentions <- bind_rows(democracy_mentions, fixed_countries) %>%
  arrange(country_name, year, Name) %>%
  mutate(has_mention = !is.na(mentions))
  
```

These constitutions have the wrong names or the wrong events:

```{r constitution_names}

democracy_mentions %>%
  ungroup() %>%
  filter(!(constitutional.event %in% c("amendment"))) %>%
  distinct(country_name, year, Name, constitutional.event, .keep_all = FALSE) %>%
  group_by(country_name, year) %>%
  filter(n() > 1) %>%
  knitr::kable()

```

So we need to fix them:

```{r constitution_names_fix}

democracy_mentions <- democracy_mentions %>%
  filter(!(country_name == "China" &
             year == 1908 &
             Name == "China_1908" &
             constitutional.event == "non-event"),
         !(country_name == "France" &
             year == 1815 &
             Name == "France's Constitution of 1815" &
             constitutional.event == "new"),
         !(country_name == "Israel" &
             year == 1948 &
             Name == "Israel's Constitution of 1948" &
             constitutional.event == "non-event"),
         !(country_name == "Myanmar (Burma)" &
             year == 2008 &
             Name == "Myanmar (Burma)'s Constitution of  2008" &
             constitutional.event == "new"),
         !(country_name == "Serbia" &
             year == 2006 &
             Name == "Yugoslavia (Serbia)'s Constitution 2006" &
             constitutional.event == "new"))
  

```

## Creation of the country-year dataset

Now we need to create the country-year dataset. We use the `tidyr` package; I assume that amendments never delete mentions of democracy.

```{r fill_in_missing}

democracy_mentions_shell <- democracy_mentions %>% 
  ungroup() %>%
  distinct(country_name, 
           GWn, 
           year,
           GW_startdate, 
           GW_enddate) %>%
  group_by(country_name, 
           GWn,
           GW_startdate,
           GW_enddate) %>%
  do(data.frame(year = min(.$year, year(.$GW_startdate[1])):ifelse(is.na(.$GW_enddate[1]), 2013, year(.$GW_enddate[1])))) %>%
  ungroup() %>%
  distinct() %>%
  select(-GW_startdate, -GW_enddate)

new_or_interim <- democracy_mentions  %>%
  filter(constitutional.event %in% c("new","interim","reinstated")) %>% 
  distinct(country_name, GWn, year, Name, constitutional.event) %>%
  right_join(democracy_mentions_shell) %>%
  arrange(country_name, year) %>%
  mutate(const_num = ifelse(!is.na(constitutional.event), 1, 0)) %>%
  group_by(country_name) %>%
  fill(Name) %>%
  mutate(const_num = cumsum(const_num)) %>%
  select(-constitutional.event) %>%
  rename(constitution_name = Name)

constitutional_events <- democracy_mentions %>%
  distinct(country_name, year, constitutional.event) 

democracy_mentions_summary <- democracy_mentions %>%
  group_by(country_name, year) %>%
  summarise(mentions = paste(unique(na.omit(mentions)), collapse = ", ")) %>%
  mutate(mentions = stringr::str_conv(mentions, "ASCII"),
         mentions = stringr::str_trunc(mentions, 500, "right"),
         mentions = ifelse(mentions == "", NA, mentions)) %>%
  ungroup()  

democracy_mentions_yearly <- democracy_mentions_shell %>%
  left_join(new_or_interim) %>%
  left_join(constitutional_events) %>%
  left_join(democracy_mentions_summary) %>%
  group_by(country_name, const_num, constitution_name) %>%
  arrange(year) %>%
  fill(mentions) %>%
  mutate(has_mention = !is.na(mentions)) %>%
  ungroup()

rm(new_or_interim,
   democracy_mentions_summary,
   democracy_mentions_shell,
   constitutional_events,
   fixed_countries,
   countries)

```

## Tests

These are Venezuela's mentions, for example:

```{r venezuela_mentions}

democracy_mentions_yearly %>% 
  filter(country_name == "Venezuela") %>% 
  group_by(country_name, constitution_name, const_num, has_mention, mentions) %>%
  summarise(constitutional.event = first(constitutional.event), min_year = min(year), max_year = max(year)) %>%
  arrange(const_num, min_year) %>%
  knitr::kable()

```

And these are Malaysia's mentions:

```{r malaysia_mentions}
democracy_mentions_yearly %>% 
  filter(country_name == "Malaysia") %>% 
  group_by(country_name, constitution_name, const_num, has_mention, mentions) %>%
  summarise(constitutional.event = first(constitutional.event), min_year = min(year), max_year = max(year)) %>%
  arrange(const_num, min_year) %>%
  knitr::kable()

```

And these are New Zealand's mentions:

```{r new_zealand_mentions}
democracy_mentions_yearly %>% 
  filter(country_name == "New Zealand") %>% 
  group_by(country_name, constitution_name, const_num, has_mention, mentions) %>%
  summarise(constitutional.event = first(constitutional.event), min_year = min(year), max_year = max(year)) %>%
  arrange(const_num, min_year) %>%
  knitr::kable()

```

And we can visualize the mentions:

```{r democracy_plot, fig.height=25}

data <- democracy_mentions_yearly %>%
  mutate(constitutional.event = fct_lump(constitutional.event, 2))

ggplot(data = data, 
       aes(x=fct_rev(reorder(country_name,year,FUN = min)), 
           y = year)) +
  geom_tile(aes(fill = has_mention)) +
  geom_point(aes(shape = constitutional.event)) +
  coord_flip() +
  labs(x = "", shape = "Constitutional\nevent", fill = "Does it mention democracy?") +
  theme_bw() +
  theme(legend.position = "top") +
  guides(fill = guide_legend(title.position = "top"),
         shape = guide_legend(title.position = "top"))

```

## Final setup

We add back the info on system membership, and make sure that when constitutions do not exist, we indicate that.

```{r final_setup}

countries <- democracy_mentions_yearly %>% 
  distinct(country_name, GWn, year) %>%
  PoliticalDatasets::to_gw_system(country_col = "country_name", 
                                  code_col = "GWn", 
                                  code_type = "GWn") %>%
  select(-country_name.x, -country_name.y) 

democracy_mentions_yearly <- countries %>%
  full_join(democracy_mentions_yearly) %>%
  mutate(has_constitution = !is.na(constitution_name)) %>%
  select(country_name, 
         GWn, 
         year, 
         constitution_name,
         const_num,
         has_constitution,
         constitutional.event,
         has_mention,
         mentions,
         GWc:in_system) 

devtools::use_data(democracy_mentions_yearly, overwrite = TRUE)
rm(countries)
```

