---
title: "Creating a latent variable index of personal power"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
bibliography: bibfile.bib
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Creating a latent variable index of personal power}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
  
This vignette shows how to create and modify the index of personal power used in several chapters of my book [*Non-democratic Politics: Authoritarianism, Dictatorships, and Democratization*](https://he.palgrave.com/page/detail/non-democratic-politics-xavier-marquez/?sf1=barcode&st1=9781137486301) (Palgrave Macmillan, 2016). It assumes that you have downloaded the replication package as follows:
  
```{r, echo = TRUE, eval = FALSE}
if(!require(devtools)) {
  install.packages("devtools")
}

devtools::install_github('xmarquez/AuthoritarianismBook')

```

It also assumes you have the `dplyr`, `ggplot2`, `scales`, `forcats`, `GGally`, `lubridate`, `reshape2`, `mirt`, and `knitr` packages installed:

```{r, echo = TRUE, eval = FALSE}
if(!require(dplyr)) {
  install.packages("dplyr")
}

if(!require(ggplot2)) {
  install.packages("ggplot2")
}

if(!require(scales)) {
  install.packages("forcats")
}

if(!require(forcats)) {
  install.packages("scales")
}

if(!require(GGally)) {
  install.packages("GGally")
}

if(!require(lubridate)) {
  install.packages("lubridate")
}

if(!require(reshape2)) {
  install.packages("reshape2")
}

if(!require(mirt)) {
  install.packages("mirt")
}

if(!require(knitr)) {
  install.packages("knitr")
}
```

```{r setup_personal_power, include=FALSE}
knitr::opts_chunk$set(fig.height = 7,
                      fig.width = 7)

```

## Preparing the data

There are three types of measures of personalism that have been used by scholars. First, there are measures of legal or de-facto "constraints on the executive;" second, there are measures of how much of a regime is "taken up" by a single executive (e.g., tenure, executive turnover); and finally, there are "all things considered" qualitative judgments of whether a regime is a "personal" one. I use all three in constructing the index of personalism used in the book.

### Indexes of constraints on the executive

The V-Dem dataset [@vdem2016dataset] contains three variables that measure constraints, legal or *de facto*, on the executive: an index of "legislative constraints on the executive" (`v2xlg_legcon`: "To what extent is the legislature and government agencies (e.g., comptroller general, general prosecutor, or ombudsman) capable of questioning, investigating, and exercising oversight over the executive?"), an index of "judicial constraints on the executive" (`v2x_jucon`: "To what extent does the executive respect the constitution and comply with court rulings, and to what extent is the judiciary able to act in an independent fashion?"), and an index of "executive corruption" (`v2x_execorr`: "How routinely do members of the executive, or their agents grant favors in exchange for bribes, kickbacks, or other material inducements, and how often do they steal, embezzle, or misappropriate public funds or other state resources for personal or family use?"). These are highly correlated, but we use all three, since they provide distinct information about the personalization of power in a regime:

```{r personal_power_data_1}

library(AuthoritarianismBook)
library(dplyr)
library(reshape2)

personalism_data <- vdem %>% 
  select(country_name, GWn, year, in_system,
         v2xlg_legcon, v2x_jucon, v2x_execorr) %>%
  melt(id.vars = c("country_name", "GWn", "year", "in_system"), na.rm = TRUE)

```

And Polity IV has a measure of constraints on the executive, the `exconst` variable (excluding the transition/interruption/interregunum codes): 

```{r personal_power_data_5}

polity_annual <- polity_annual %>%  mutate(exconst = ifelse(unclass(exconst) < 4, NA, unclass(exconst) - 3))

personalism_data <- bind_rows(personalism_data, polity_annual %>%
                                select(country_name, GWn, year, exconst, in_system) %>%
                                melt(id.vars = c("country_name","GWn","year", "in_system"))) 

rm(polity_annual)
```

```{r correlations_vdem_personal}

library(ggplot2)
library(GGally)
library(reshape2)

ggcorr(dcast(personalism_data, GWn + year + in_system ~ variable) %>% 
         select(-GWn, -in_system, -year), label = TRUE,
  label_round = 2)

```

These measures are generally reasonable. Here's the graph for Russia, for example:

```{r Russia_constraints_graph}

library(lubridate)

leader_data_yearly <- archigos %>% 
  group_by(obsid, GWn, leader, startdate,enddate) %>%
  do(data.frame(year = year(.$startdate):year(.$enddate))) %>%
  group_by(GWn, year) %>%
  filter(enddate == max(enddate), startdate == max(startdate)) %>%
  group_by(leader) %>%
  mutate(is_start_leader_year = (year == min(year))) %>%
  ungroup()

country <- "Russia (Soviet Union)"
  
data <- left_join(personalism_data, leader_data_yearly) %>%
  filter(country_name %in% country)

extra_breaks <- NULL

extra_labels <- NULL

axis_labels <- data_frame(breaks = c(data$year[data$is_start_leader_year], extra_breaks),
                          labels = c(paste(data$year[data$is_start_leader_year],
                                    data$leader[data$is_start_leader_year]), extra_labels)) %>%
  distinct() %>%
  group_by(breaks) %>%
  summarise(labels = paste(labels, collapse = "\n"))


ggplot(data = data %>% 
         filter(grepl("v2", variable)), 
       aes(x=year, y = value,
           color = variable,
           fill = variable)) +
    geom_path() +
    facet_wrap(~country_name, ncol=2) +
    theme_bw() +
    scale_x_continuous(breaks = axis_labels$breaks,
                       labels = axis_labels$labels) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_vline(data = data %>%
                 filter(is_start_leader_year) %>%
                 distinct(leader,value, is_start_leader_year, .keep_all = TRUE),
               aes(xintercept = year)) +
    labs(y = "Degree of constraint/corruption
         (in standard deviations from mean, 
         0 is the mean in sample)", x = "Year") +
    theme(legend.position = "bottom") +
    guides(linetype = guide_legend(title = ""),
           color = guide_legend(title = ""),
           fill = guide_legend(title = ""))

ggplot(data = data %>% 
         filter(grepl("exconst", variable)), 
       aes(x=year, y = value,
           color = variable,
           fill = variable)) +
    geom_path() +
    facet_wrap(~country_name, ncol=2) +
    theme_bw() +
    scale_x_continuous(breaks = axis_labels$breaks,
                       labels = axis_labels$labels) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_vline(data = data %>%
                 filter(is_start_leader_year) %>%
                 distinct(leader,value, is_start_leader_year, .keep_all = TRUE),
               aes(xintercept = year)) +
    labs(y = "Degree of constraint/corruption
         (in standard deviations from mean, 
         0 is the mean in sample)", x = "Year") +
    theme(legend.position = "bottom") +
    guides(linetype = guide_legend(title = ""),
           color = guide_legend(title = ""),
           fill = guide_legend(title = ""))

```

The graphs basically say that during Soviet times, legislative and executive constraints on the executive were greatly weakened. Executive corruption was at the mean in the sample, though it increased with Brezhnev (as we might have guessed from case knowledge), and it really shot upwards with the break up of the Soviet Union. At the same time, legislative and executive constraints increased until Putin came into office, when they began to decrease again.

### Tenure-based measures of executive constraint

The dataset of regime types by @MagaloniChuMin2013 contains two measures of personalism: one derived from the `exconst` variable of the Polity IV dataset, and another one (`lindex`) which they describe as follows:

> the variable is essentially a Herfindahl index (sum of squared shares) using the name of the executive. For a given country-year in a unique regime (see reg_id), the following calculation is made: $\sum_{i=1}^m(exec_i/n)^2$ where n is the age of the regime up to that year, and exec is the number of years that a unique executive i (out of a total m executives up to that year) has led the regime. As such, a regime led by only one person up through that year yields a personalism index of 1. A theoretical scenario where leadership changes every single year would yield $1/n$. These calculations are made using the non-rounded values. We note that this is a relatively sensitive measure in the early/formative years of an individual regime, but we propose this is a useful way of considering personalism as an evolving attribute of a regime over time. 

I use this `lindex` variable only (since I directly included the `exconst` variable, noted above):

```{r personal_power_data_2}

personalism_data <- bind_rows(personalism_data, magaloni %>%
                                select(country_name, GWn, year, in_system, lindex) %>%
                                melt(id.vars = c("country_name", "GWn", "year", "in_system"), na.rm = TRUE)) 
```

Wahman, Teorell, and Hadenius [@WahmanTeorellHadenius2013] also propose a measure of personalism based on "mean executive turnover": the number of "changes of the chief executive during the regime spell, divided by the number of years of regime spell duration", according" according to their regime classification. In a highly personalized regime this should be zero, while it ranges up to 3 chief executives per year for regimes with significant personnel turnover:

```{r personal_power_data_3}

personalism_data <- bind_rows(personalism_data, wahman_teorell %>%
                                select(country_name, GWn, year, persagg1ny, in_system) %>%
                                melt(id.vars = c("country_name", "GWn", "year", "in_system"), na.rm= TRUE)) 

```

These measures are not well correlated with the measures of executive constraint or with one another. The `lindex` measure is also always 1 during the first year of the regime, which ensures it will always show "high" personalism at the time. This is perhaps justified as capturing the fact that new regimes are not yet highly institutionalized; every regime is highly personalized in its first year, until a regular transfer of power happens. But it is unintuitive, seemingly suggesting that both democratic and non-democratic regimes are equally personalized early on. At any rate, the lack of correlation between these measures indicates that they are at best measuring a different dimension of the phenomenon of personal rule than the constraints-based measures: 

```{r tenure_measures_corr}

ggcorr(dcast(personalism_data, GWn + year + in_system ~ variable) %>% 
         select(-GWn:-in_system), label = TRUE,
  label_round = 2)

```

It is also worth noting that both `lindex` and `persagg1ny` are highly sensitive to the particular periodization of regimes used. For example, using Archigos [@GoemansGleditschChiozza2009], we can reconstruct both the @WahmanTeorellHadenius2013 and the @MagaloniChuMin2013 indexes for the @GeddesWrightFrantz2014 classification of political regimes:

```{r reconstruction_persagg1ny}

cumulative_levels <- function(seq) {
  seq <- as.factor(seq)
  levels(seq) <- seq_along(levels(seq))
  as.numeric(seq)
}

reconstruction <- magaloni %>%
  select(country_name, GWn, year, in_system, lindex) %>%
  full_join(magaloni_extended %>% 
              select(country_name, GWn, year, in_system, regime_nr)) %>%
  full_join(wahman_teorell %>% 
              select(country_name, GWn, year, in_system, regime1ny, persagg1ny)) %>%
  full_join(all_gwf_extended_yearly %>%
              select(country_name, GWn, year, in_system, gwf_casename, gwf_full_regimetype)) %>%
  full_join(leader_data_yearly) %>%
  arrange(country_name, year, startdate) %>%
  filter(!is.na(obsid), !is.na(gwf_casename)) %>%
  group_by(gwf_casename) %>%
  mutate(gwf_persagg1ny = (length(unique(obsid)) - 1)/length(unique(year))) %>%
  ungroup()

ggcorr(reconstruction %>% 
         select(gwf_persagg1ny, persagg1ny, lindex), 
       label = TRUE,
       label_round = 2)
```

```{r lindex_reconstruction}

lindex_reconstruction <- reconstruction %>%
  group_by(gwf_casename) %>%
  arrange(year) %>%
  mutate(m = cumulative_levels(obsid),
         m = m - min(m) + 1,
         max_year = max(year)) %>%
  group_by(gwf_casename, obsid) %>%
  mutate(min_year = min(year),
         max_tenure = n()) %>%
  group_by(country_name, gwf_casename, leader, m, max_tenure, obsid) %>%
  do(data.frame(year = unique(.$min_year):unique(.$max_year))) %>%
  mutate(exec_i = seq_along(year),
         exec_i = ifelse(exec_i > max_tenure, 
                         max_tenure,
                         exec_i)) %>%
  group_by(gwf_casename) %>%
  arrange(year) %>%
  mutate(n = seq_along(year),
         share = (exec_i/n)^2) %>%
  group_by(country_name, gwf_casename, year) %>%
  summarise(gwf_lindex = sum(share))

reconstruction <- left_join(reconstruction, lindex_reconstruction)

ggcorr(reconstruction %>% 
         select(gwf_persagg1ny, persagg1ny, lindex, gwf_lindex), 
       label = TRUE,
       label_round = 2)


```

The correlations between the original and reconstructed variables are only moderately strong, since the regime periodizations in the respective datasets of political regimes are not identical. For example, here is what original and reconstructed variables look for Russia:

```{r lindex_persagg1ny}

data <- reconstruction %>% 
  filter(country_name %in% country) %>%
  melt(measure.vars = c("gwf_persagg1ny", "persagg1ny", "lindex", "gwf_lindex"))

ggplot(data = data %>% filter(grepl("persa", variable)), 
       aes(x=year, y = value,
           color = variable,
           fill = variable)) +
    geom_path() +
    facet_wrap(~country_name, ncol=2) +
    theme_bw() +
    scale_x_continuous(breaks = axis_labels$breaks,
                       labels = axis_labels$labels) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_vline(data = data %>%
                 filter(is_start_leader_year) %>%
                 distinct(leader,value, is_start_leader_year, .keep_all = TRUE),
               aes(xintercept = year)) +
    labs(y = "Persagg1ny", x = "Year") +
    theme(legend.position = "bottom") +
    guides(linetype = guide_legend(title = ""),
           color = guide_legend(title = ""),
           fill = guide_legend(title = ""))

ggplot(data = data %>% filter(grepl("lindex", variable)), 
       aes(x=year, y = value,
           color = variable,
           fill = variable)) +
    geom_path() +
    facet_wrap(~country_name, ncol=2) +
    theme_bw() +
    scale_x_continuous(breaks = axis_labels$breaks,
                       labels = axis_labels$labels) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_vline(data = data %>%
                 filter(is_start_leader_year) %>%
                 distinct(leader,value, is_start_leader_year, .keep_all = TRUE),
               aes(xintercept = year)) +
    labs(y = "Lindex", x = "Year") +
    theme(legend.position = "bottom") +
    guides(linetype = guide_legend(title = ""),
           color = guide_legend(title = ""),
           fill = guide_legend(title = ""))
```

Because `persagg1ny` is invariant over the life of a regime, it is insensitive to fluctuations in personal power; by this measure, Lenin, Stalin, and Brezhnev all have the same amount of personal power over the life of the Soviet regime, since the "personalism" of the regime is averaged throughout its entire duration. Similarly, because `lindex` is always 1 at the beginning of a regime, the measure barely begins to capture Stalin's personal power as his tenure lengthened. Despite these problems, we nevertheless include both `lindex` and `persagg1ny` in the construction of the personal power index; in any case the model will discount them when they are highly uncorrelated with the rest of the other measures.

```{r correlations_with_rest}

personalism_data <- bind_rows(personalism_data, reconstruction %>%
                                select(country_name, GWn, year, in_system, gwf_lindex, gwf_persagg1ny) %>%
                                melt(id.vars = c("country_name","GWn","year", "in_system")))

personalism_data <- personalism_data %>% 
  filter(!is.na(value)) %>% 
  distinct()

ggcorr(dcast(personalism_data, country_name + GWn + year + in_system ~ variable) %>% 
         select(-country_name:-in_system), label = TRUE,
  label_round = 2)

```


### All-things-considered measures of personal rule

Finally, two datasets contain "all things considered" measures of whether a regime is "personal" (or has a personalistic component): @Kailitz2013 and @GeddesWrightFrantz2014. 

```{r personal_power_data_4}

personalism_data <- bind_rows(personalism_data, kailitz_yearly %>%
                                select(country_name, GWn, year, in_system, personal) %>%
                                melt(id.vars = c("country_name", "GWn", "year", "in_system"))) 
```

```{r personal_power_data_6}

all_gwf_extended_yearly <- all_gwf_extended_yearly %>%
  select(country_name, GWn, year, in_system, gwf_full_regimetype) %>%
  mutate(gwf_personal_single = grepl("^personal",gwf_full_regimetype),
         gwf_personal_hybrid  = grepl("personal",gwf_full_regimetype),
         gwf_personal = gwf_personal_single + gwf_personal_hybrid) %>%
  select(-gwf_personal_single,-gwf_personal_hybrid)

personalism_data <- bind_rows(personalism_data, all_gwf_extended_yearly %>%
                                select(-gwf_full_regimetype) %>%
                                melt(id.vars = c("country_name", "GWn", "year", "in_system"))) 

rm(all_gwf_extended_yearly)
```

Both of these measures represent coarse qualitative judgments, and are not perfectly correlated:

```{r personal_correlations}

ggcorr(dcast(personalism_data, country_name + GWn + year + in_system ~ variable) %>% 
         select(gwf_personal, personal), label = TRUE,
  label_round = 2)

```

Nevertheless, both add some information to our index of personal power, and are therefore both used. The total correlation matrix looks like this:

```{r correlations_1}

ggcorr(data = dcast(personalism_data, GWn + year + in_system ~ variable) %>% 
         select(exconst:v2xlg_legcon),
       label = TRUE,
       label_round = 2)

```

## Creating the index 

### Cutting the data into a a manageable number of intervals

Before running the model, it is necessary to cut the continuous variables into a finite number of intervals. For entirely arbitrary reasons, I use 20 intervals. I also first transform the `lindex` and `persagg1ny` measures into percentile ranks. In the case of `lindex`, this smooths out some of the spikes of personal power at the beginning of a regime.

```{r personal_power_data_7}

personalism_data <- dcast(personalism_data, country_name + GWn + year + in_system ~ variable)

summary(personalism_data)
```


```{r personal_power_data_9}

personalism_data <- personalism_data %>% 
  mutate_at(.funs = funs(unclass(cut(., 
                               breaks = 20, 
                               include.lowest = TRUE, 
                               right = FALSE, 
                               ordered_result = TRUE))),
              .cols = c("v2x_execorr", "v2xlg_legcon", "v2x_jucon"))

personalism_data <- personalism_data %>% 
  mutate_at(.funs = funs(unclass(cut(percent_rank(.), 
                              20, 
                              include.lowest=TRUE, 
                              ordered_result=TRUE))),
              .cols = c("lindex", "gwf_lindex", "persagg1ny", "gwf_persagg1ny"))

personalism_data <- personalism_data %>% 
  mutate_at(.funs = funs(as.numeric(unclass(factor(., 
                                  ordered = TRUE)))), 
            .cols = c("v2x_execorr", "v2xlg_legcon", "v2x_jucon", "exconst",
                   "lindex", "gwf_lindex", "persagg1ny", "gwf_persagg1ny",
                   "personal", "gwf_personal"))
```

With these cuts, here is what the measures look like for Russia:

```{r check_Russia}

data <- left_join(personalism_data, leader_data_yearly) %>%
  filter(country_name %in% country) %>%
  melt(measure.vars = c("lindex",
                        "gwf_lindex",
                        "persagg1ny",
                        "gwf_persagg1ny",
                        "v2xlg_legcon", 
                        "v2x_jucon", 
                        "v2x_execorr",
                        "exconst"))

ggplot(data = data, 
       aes(x=year, y = value,
           color = variable,
           fill = variable)) +
    geom_path() +
    facet_wrap(~country_name, ncol=2) +
    theme_bw() +
    scale_x_continuous(breaks = axis_labels$breaks,
                       labels = axis_labels$labels) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_vline(data = data %>%
                 filter(is_start_leader_year) %>%
                 distinct(leader,value, is_start_leader_year, .keep_all = TRUE),
               aes(xintercept = year)) +
    labs(y = "Transformed variables", x = "Year") +
    theme(legend.position = "bottom") +
    guides(linetype = guide_legend(title = ""),
           color = guide_legend(title = ""),
           fill = guide_legend(title = ""))

```

The correlations between the transformed variables are very similar to the correlations between the untransformed variables:

```{r correlations_2}

library(GGally)

ggcorr(data= personalism_data %>% select(exconst:v2xlg_legcon), label = TRUE, label_round = 2)

```

### Running the model

Finally we create the model. Model 1 is the one used in the book:

```{r model_1, eval = TRUE}

data_1 <- personalism_data %>% 
  melt(measure.vars = c("exconst",
                        "gwf_personal",
                        "lindex", 
                        "persagg1ny",
                        "personal",
                        "v2x_execorr",
                        "v2x_jucon",
                        "v2xlg_legcon"),
       na.rm = TRUE) %>%
  dcast(country_name + GWn + year + in_system ~ variable)
  

library(mirt)
personal_model <- mirt(data_1 %>% 
                         select(exconst, 
                                gwf_personal,
                                lindex, 
                                persagg1ny,
                                personal,
                                v2x_execorr, 
                                v2x_jucon, 
                                v2xlg_legcon), 
                   model = 1, 
                   itemtype = "graded", 
                   SE = TRUE, 
                   verbose = FALSE)

personal_model

summary(personal_model)
```

In model 1, the latent factor is most closely correlated with the measures of executive constraint, though Kailitz and Geddes, Wright, and Frantz's measures of personalism get some weight too. 

Model 2 uses instead the constructed measures of `lindex` and `persagg1ny` created from GWF's regime classification and the ARCHIGOS data:

```{r model_2, eval = TRUE }

data_2 <- personalism_data %>% 
  melt(measure.vars = c("exconst",
                        "gwf_personal",
                        "gwf_lindex", 
                        "gwf_persagg1ny",
                        "personal",
                        "v2x_execorr",
                        "v2x_jucon",
                        "v2xlg_legcon"),
       na.rm = TRUE) %>%
  dcast(country_name + GWn + year + in_system ~ variable)

personal_model_2 <- mirt(data_2 %>% 
                         select(v2x_execorr, v2xlg_legcon, v2x_jucon, exconst,
                   gwf_lindex, gwf_persagg1ny, personal, gwf_personal), 
                   model = 1, 
                   itemtype = "graded", 
                   SE = TRUE, 
                   verbose = FALSE)

personal_model_2

summary(personal_model_2)
```

Model 2 puts a lot of weight on the GWF-based `persagg1ny` measure. 

Model 3 excludes the tenure-based measures of personal power completely:

```{r model_3, eval = TRUE }

data_3 <- personalism_data %>% 
  melt(measure.vars = c("exconst",
                        "gwf_personal",
                        "personal",
                        "v2x_execorr",
                        "v2x_jucon",
                        "v2xlg_legcon"),
       na.rm = TRUE) %>%
  dcast(country_name + GWn + year + in_system ~ variable)

personal_model_3 <- mirt(data_3 %>% 
                         select(v2x_execorr, v2xlg_legcon, v2x_jucon, exconst,
                   personal, gwf_personal), 
                   model = 1, 
                   itemtype = "graded", 
                   SE = TRUE, 
                   verbose = FALSE)

personal_model_3

summary(personal_model_3)
```

```{r join_all_model_scores, eval = TRUE}
personal_scores <- fscores(personal_model, full.scores = TRUE, full.scores.SE = TRUE) %>% 
  data.frame() %>% 
  dplyr::rename(z1 = F1, se.z1 = SE_F1) %>% 
  mutate(pct975 = z1 + 1.96 * se.z1, pct025 = z1 - 1.96 * se.z1)

personal_scores <- bind_cols(data_1 %>% select(country_name, GWn, year), personal_scores)

personal_scores_2 <- fscores(personal_model_2, full.scores = TRUE, full.scores.SE = TRUE) %>% 
  data.frame() %>% 
  dplyr::rename(z1 = F1, se.z1 = SE_F1) %>% 
  mutate(pct975 = z1 + 1.96 * se.z1, pct025 = z1 - 1.96 * se.z1)

personal_scores_2 <- bind_cols(data_2 %>% select(country_name, GWn, year), personal_scores_2)

personal_scores_3 <- fscores(personal_model_3, full.scores = TRUE, full.scores.SE = TRUE) %>% 
  data.frame() %>% 
  dplyr::rename(z1 = F1, se.z1 = SE_F1) %>% 
  mutate(pct975 = z1 + 1.96 * se.z1, pct025 = z1 - 1.96 * se.z1)

personal_scores_3 <- bind_cols(data_3 %>% select(country_name, GWn, year), personal_scores_3)

personal_scores <- bind_rows(personal_scores, 
                             personal_scores_2, 
                             personal_scores_3, 
                             .id = "model")

rm(data_1, data_2, personal_scores_2, personal_scores_3)

```

The index right now is inverted - shows higher values of personal power as lower numbers - so we typically want to reverse it:

```{r reverse_index, eval = TRUE}

personal_scores <- personal_scores %>%
  mutate(z1 = -z1,
         pct975 = -pct975,
         pct025 = -pct025)

```

With that adjustment, this is what these scores look like for Russia:

```{r Russia_test_personal_scores, eval = TRUE}

data <- left_join(personal_scores, leader_data_yearly) %>%
  filter(country_name %in% country)

ggplot(data = data, 
       aes(x=year, y = z1,
           ymin = pct025,
           ymax = pct975,
           color = model,
           fill = model)) +
  geom_path() +
  geom_ribbon(alpha = 0.3) +
  facet_wrap(~country_name, ncol=2) +
  theme_bw() +
  scale_x_continuous(breaks = axis_labels$breaks,
                     labels = axis_labels$labels) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  geom_vline(data = data %>%
               filter(is_start_leader_year) %>%
               distinct(leader, is_start_leader_year, .keep_all = TRUE),
             aes(xintercept = year)) +
  labs(y = "Index score", x = "Year") +
  theme(legend.position = "bottom") +
  guides(linetype = guide_legend(title = ""),
         color = guide_legend(title = "Model"),
         fill = guide_legend(title = "Model"))

```

The scores from each of these models are in any case extremely well correlated:

```{r scores_corr, eval = TRUE}

data <- personal_scores %>%
  mutate(model = plyr::mapvalues(model, 
                                 from = c(1,2,3),
                                 to = c("Model_1_in_book","Model_2","Model_3"))) %>%
  dcast(country_name + GWn + year ~ model, value.var = "z1") %>%
  select(Model_1_in_book:Model_3)

ggcorr(data = data, label = TRUE, label_round = 2)

```

## Full set of scores

We can visualize the full set of scores (this takes a long time to run, so it is not run by default; set `eval = TRUE` to run the chunk when compiling the vignette):

```{r visual_test, eval = FALSE }

library(lubridate)

# Visual test
leader_data <- archigos %>% 
  group_by(GWn, leader, startdate, enddate) %>%
  summarise(year = year(startdate)) %>%
  group_by(GWn, year) %>%
  filter(enddate == max(enddate), 
         startdate == max(startdate))


personal_data <- left_join(personal_scores,leader_data) %>%
  group_by(country_name, model) %>%
  arrange(model, country_name, year) %>%
  ungroup()
  

for(country in unique(personal_data$country_name)) {

  data <- personal_data %>% filter(country_name %in% country, year > 1900)
  
  if(length(data$year[!is.na(data$leader)]) > 0) {
  p <- ggplot(data = data, aes(x=year,
                               y = z1, 
                               ymin = pct025, 
                               ymax = pct975,
                               color = model,
                               fill = model)) +
    geom_path() +
    geom_ribbon(alpha=0.2) +
    theme_bw() +
    scale_x_continuous(breaks = data$year[!is.na(data$leader) & data$model == 1 ],
                       labels = paste(data$year[!is.na(data$leader) & data$model == 1 ], 
                                      data$leader[!is.na(data$leader) & data$model == 1 ])) +
    theme(axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     vjust = 0.5,
                                     size = 5)) +
    labs(x = "") +
    facet_wrap(~country_name, scales="free")
  
  print(p)
    
  }

}


```

## References
