---
title: "Chapter 6: Armies"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
bibliography: bibfile.bib
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Chapter 6}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
## Introduction
  
This vignette shows how to replicate and extend the charts in chapter 6 of my book [*Non-democratic Politics: Authoritarianism, Dictatorships, and Democratization*](https://he.palgrave.com/page/detail/non-democratic-politics-xavier-marquez/?sf1=barcode&st1=9781137486301) (Palgrave Macmillan, 2016). It assumes that you have downloaded the replication package as follows:
  
```{r, echo = TRUE, eval = FALSE}
if(!require(devtools)) {
  install.packages("devtools")
}

devtools::install_github('xmarquez/AuthoritarianismBook')

```

It also assumes you have the `dplyr`, `ggplot2`, `scales`, `forcats`, and `knitr` packages installed:

```{r, echo = TRUE, eval = FALSE}
if(!require(dplyr)) {
  install.packages("dplyr")
}

if(!require(ggplot2)) {
  install.packages("ggplot2")
}

if(!require(scales)) {
  install.packages("forcats")
}

if(!require(forcats)) {
  install.packages("scales")
}

if(!require(knitr)) {
  install.packages("knitr")
}
```

```{r setup_chapter_6, include=FALSE}
knitr::opts_chunk$set(fig.height = 7, 
fig.width = 7)
```

## Figure 6.1: Successful and unsuccessful military coups, 1950-2015

### Original figure

Successful and unsuccessful military coups, 1950-2015. Data from @PowellThyne2011, updated to 2015. The data used to construct these figures slightly overstate the frequency of genuine military coups, since they include coups led by non-military insiders.

```{r fig_6_1}
library(AuthoritarianismBook)
library(dplyr)
library(ggplot2)

data <- powell_thyne

ggplot(data=data, aes(x=year, fill = attempt_type)) + 
  geom_bar() +
  theme_bw() +
  labs(y = "Number of coups", fill = "Coup success") + 
  guides(fill = guide_legend(title.position = "top")) +
  theme(legend.position="bottom")

```

### Extension: Map of countries by number of coups

We can also visualize more or less coup-prone countries:

```{r coup_prone}

world <- world %>% filter(id != "Antarctica")

data <- data %>% 
  mutate(country_name = ifelse(country_name == "Germany (Prussia)", "German Federal Republic", country_name)) %>%
  count(country_name)

ggplot() + 
  geom_map(aes(fill = n, map_id = country_name), map = world, data = data) +
  geom_path(data = world,aes(y = lat,x = long,group = group)) +
  scale_fill_gradient2() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(fill = "Total number of successful and\nunsuccessful coups since 1950",x = "",y = "") +
  guides(fill = guide_legend(title.position = "top")) +
  scale_y_continuous(breaks=NULL) +
  scale_x_continuous(breaks=NULL) 

```

## Figure 6.2: Level of democracy before and after successful coups, 1950-2015

This figure uses both @PowellThyne2011 and the extended Unified Democracy Scores [@Pemstein2010; @Marquez2016]. The smooth solid black line below represents a smoothed line of fit for the average level of democracy in a country before a successful coup took place (lowess smooth); the dashed line represents a smoothed line of fit for the level of democracy after the coup took place. Each small gray arrow represents the change in the measured level of democracy in a given country produced by a successful coup. Scores above 0.5 can be considered more or less democratic. Chile's democracy scores for the entire period are provided for illustrative purposes. 

```{r fig_6_2}

data <- left_join(extended_uds, powell_thyne %>% 
                    filter(attempt_type == "Successful") %>%
                    group_by(country_name, year) %>% 
                    summarise(total_coups = n())) %>% 
  group_by(country_name) %>%
  mutate(lagged_index = lag(index)) %>%
  filter(year >= 1950) 

ggplot(data= data %>% filter(!is.na(total_coups))) +
  geom_segment(aes(x = year-1, 
                   xend = year,
                   y = lagged_index,
                   yend = index),
               alpha = 0.2,
               arrow = arrow(length = unit(0.1, "cm"))) +
  geom_smooth(aes(x= year,y = index),
              color="red",
              linetype=2) +
  geom_smooth(aes(x = year,y= lagged_index), 
              color="blue",
              linetype=1) +
  geom_path(data= data %>% filter(country_name == "Chile"),
            aes(x = year,
                y = lagged_index)) +
  geom_text(data = data %>% filter(!is.na(total_coups), country_name == "Chile"), 
            aes(x = year,
                y = lagged_index,
                label = paste(country_name,year))) +
  theme_bw() +
  geom_hline(yintercept=0.5) +
  labs(x = "Year", 
       y = "Unified democracy score (democracy > 0.5)")

```

## Figure 6.3: Estimated duration of regimes with substantial military participation

This figure assumes you have installed the packages `survival`, `broom`, and `lubridate`. You can install them as follows:

```{r, echo = TRUE, eval = FALSE}
if(!require(survival)) {
  install.packages("survival")
}

if(!require(broom)) {
  install.packages("broom")
}

if(!require(survival)) {
  install.packages("survival")
}

```

The figure models the survival of regimes without any correlates other than regime type. It uses the data from @GeddesWrightFrantz2014 to measure the survival of regimes. The figure here is slightly different from the published figure since; it corrects a small error.

```{r fig_6_3}

library(broom)
library(survival)
library(lubridate)

data <- all_gwf_periods %>% 
  mutate(gwf_fail = (gwf_enddate != ymd("2010-12-31")), 
         military_regime = ifelse(grepl("military",
                                        gwf_full_regimetype,
                                        ignore.case=TRUE),
                                  "Military",
                                  ifelse(grepl("democracy",
                                               gwf_full_regimetype,
                                               ignore.case=TRUE),
                                         "Democracy",
                                         "Other non-democratic")))

survival.model <- survfit(Surv(time = as.double(gwf_enddate - gwf_startdate,
                                                units="days"),
                               event = gwf_fail) ~ military_regime, 
                          data = data) %>%
  tidy() %>% 
  mutate(strata = gsub("military_regime=","", x = strata))

survival.model$strata <- forcats::fct_relevel(as.factor(survival.model$strata),
                                              "Military", 
                                              "Other non-democratic")

ggplot(data = survival.model,
       aes(x = time/365, y = estimate,
           linetype = strata, fill = strata)) + 
  geom_line() +
  theme_bw() +
  labs(y = "Proportion of regimes surviving",
       x = "Regime duration in years",
       fill = "Regime type", 
       linetype = "Regime type") + 
  theme(legend.position="bottom")+
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.3) + 
  coord_cartesian(xlim = c(0, 58)) +
  scale_y_continuous(labels = scales::percent)

```

This model can be estimated by omitting regimes that are neither democratic nor non-democratic (state failure, etc.):

```{r fig_6_3_published }

data <- all_gwf_periods %>% 
  filter(!(gwf_full_regimetype %in% c("warlord",
                                      "warlord/foreign-occupied",
                                      "provisional",
                                      "foreign-occupied",
                                       "not-independent"))) %>%
  mutate(gwf_fail = (gwf_enddate != ymd("2010-12-31")), 
         military_regime = ifelse(grepl("military",
                                        gwf_full_regimetype,
                                        ignore.case=TRUE),
                                  "Military",
                                  ifelse(grepl("democracy",
                                               gwf_full_regimetype,
                                               ignore.case=TRUE),
                                         "Democracy",
                                         "Other non-democratic")))

survival.model <- survfit(Surv(time = as.double(gwf_enddate - gwf_startdate,
                                                units="days"),
                               event = gwf_fail) ~ military_regime, 
                          data = data) %>%
  tidy() %>% 
  mutate(strata = gsub("military_regime=","", x = strata))

survival.model$strata <- forcats::fct_relevel(as.factor(survival.model$strata),
                                              "Military", 
                                              "Other non-democratic")

ggplot(data = survival.model,
       aes(x = time/365, y = estimate,
           linetype = strata, fill = strata)) + 
  geom_line() +
  theme_bw() +
  labs(y = "Proportion of regimes surviving",
       x = "Regime duration in years",
       fill = "Regime type", 
       linetype = "Regime type") + 
  theme(legend.position="bottom")+
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.3) + 
  coord_cartesian(xlim = c(0, 58)) +
  scale_y_continuous(labels = scales::percent)
```

## Figure 6.4: Estimated duration of non-democratic regimes by type of military involvement

Here we use data on military involvement in non-democratic regimes collected by @Svolik2012.

```{r fig_6_4}

library(broom)
library(survival)

count_sequence_breaks <- function(seq, seq_step = 1) {
  first_diff <- c(seq_step, diff(seq)) - seq_step
  periods <- cumsum(abs(first_diff))
  periods
}

data <- svolik_institutions %>%
  group_by(country_name) %>%
  mutate(military = ifelse(is.na(military),"Not available", military),
         mil_regime_no = count_sequence_breaks(unclass(as.factor(military)), seq_step=0),
         svolik_fail = (year != max(year))) %>%
  group_by(country_name, mil_regime_no, military) %>%
  summarise(min_year = min(year), max_year = max(year), svolik_fail = min(svolik_fail)) %>%
  filter(military != "Not available") %>%
  mutate(military = as.factor(military),
         military = forcats::fct_recode(military, 
                                        "Civilian (no military involvement)" = "civilian"))

survival.model <- survfit(Surv(time = max_year - min_year,
                               event = svolik_fail) ~ military, 
                          data=data) %>%
  tidy() %>%
  mutate(strata=gsub("military=","",x=strata))

ggplot(data = survival.model, 
       aes(x = time,y = estimate, linetype = strata, fill = strata)) +
  geom_line() +
  theme_bw() +
  labs(y = "Proportion of non-democratic regimes surviving",
       x = "Regime duration in years",
       fill = "Type of military involvement",
       linetype = "Type of military involvement") +
  theme(legend.position = "bottom") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  scale_y_continuous(labels = scales::percent)

```

## References

