---
title: "Chapter 10: The Roots of Regime Change and Democratization"
author: "Xavier Marquez"
date: "22 December 2016"
bibliography: bibfile.bib
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Chapter 10}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
# Introduction
  
This vignette shows how to replicate and extend the charts in chapter 10 of my book [*Non-democratic Politics: Authoritarianism, Dictatorships, and Democratization*](https://he.palgrave.com/page/detail/non-democratic-politics-xavier-marquez/?sf1=barcode&st1=9781137486301) (Palgrave Macmillan, 2016). It assumes that you have downloaded the replication package as follows:
  
```{r, echo = TRUE, eval = FALSE}
if(!require(devtools)) {
  install.packages("devtools")
}

devtools::install_github('xmarquez/AuthoritarianismBook')

```

It also assumes you have the `dplyr`, `ggplot2`, `scales`, `forcats`, `broom`, `reshape2`, `tidyr`, `purrr`, and `knitr` packages installed:

```{r, echo = TRUE, eval = FALSE}
if(!require(dplyr)) {
  install.packages("dplyr")
}

if(!require(ggplot2)) {
  install.packages("ggplot2")
}

if(!require(scales)) {
  install.packages("forcats")
}

if(!require(forcats)) {
  install.packages("scales")
}

if(!require(broom)) {
  install.packages("broom")
}

if(!require(reshape2)) {
  install.packages("reshape2")
}

if(!require(purrr)) {
  install.packages("purrr")
}

if(!require(tidyr)) {
  install.packages("tidyr")
}

if(!require(knitr)) {
  install.packages("knitr")
}
```

For the interactive graph below, you also need the `plotly` package:

```{r, echo = TRUE, eval = FALSE}
if(!require(plotly)) {
  install.packages("plotly")
}
```

```{r setup_chapter_10, include=FALSE}
knitr::opts_chunk$set(fig.height = 7, 
fig.width = 7)
```

# Table 10.1: Frequency of different types of regime transition, 1945-2010

## Original Table

This table shows the frequency of different types of regime transition, 1945-2010. This table counts only countries that have experienced transitions. Regime change is rare; in the period 1945-2010, only about 5% of all country-years experienced regime transitions. Regime data comes from @GeddesWrightFrantz2014. `Other' category includes instances of state breakdown, foreign occupation, and state termination (e.g., after unification with a larger state).

```{r table_10_1}

library(AuthoritarianismBook)
library(dplyr)
library(ggplot2)
library(reshape2)

data <- all_gwf_periods %>% 
  mutate(gwf_previous = ifelse(grepl("democracy",
                                     gwf_full_regimetype),
                               "Democracy",
                               ifelse(grepl("military|oligarchy|personal|party|monarchy",
                                            gwf_full_regimetype),
                                      "Non-democracy",
                                      "Other"))) %>%
  group_by(country_name) %>%
  mutate(gwf_next = lead(gwf_previous)) %>%
  ungroup() %>%
  count(gwf_previous, gwf_next, sort=TRUE) %>% 
  filter(!is.na(gwf_next)) %>%
  ungroup() %>%
  mutate(total_trans = sum(n), 
         prop_trans = scales::percent_format()(n/total_trans)) 

data  %>% 
  dcast(gwf_previous ~ gwf_next, value.var="prop_trans") %>%
  knitr::kable(col.names = c("Previous regime", "To democracy", "To non-democracy", "To other"))



```

## Extension 1: Total number of transitions

This table simply shows the total number of transitions, rather than the proportions in the datatset:

```{r table_10_1_total_n}

data  %>% 
  dcast(gwf_previous ~ gwf_next, value.var="n") %>%
  knitr::kable(col.names = c("Previous regime", "To democracy", "To non-democracy", "To other"),
               caption = "Total number of transition events in the dataset")

```

## Extension 2: Proportions per row and column

Here we see the proportions per row:

```{r table_10_1_proportions_per_row}

data %>%
  group_by(gwf_previous) %>%
  mutate(total_by_row = sum(n), prop_trans = scales::percent_format()(n/total_by_row)) %>% 
  dcast(gwf_previous ~ gwf_next, value.var="prop_trans") %>%
  knitr::kable(col.names = c("Previous regime", "To democracy", "To non-democracy", "To other"),
               caption = "Proportions by row (previous regime)")
```

More than 98 per cent of transitions from democracy are to non-democracy; only 36 per cent of transitions from non-democracy are to democracy, and nearly 50 per cent are to other forms of non-democracy.

```{r table_10_1_proportions_per_column}

data %>%
  group_by(gwf_next) %>%
  mutate(total_by_column = sum(n), prop_trans = scales::percent_format()(n/total_by_column)) %>% 
  dcast(gwf_previous ~ gwf_next, value.var="prop_trans") %>%
  knitr::kable(col.names = c("Previous regime", "To democracy", "To non-democracy", "To other"),
               caption = "Proportions by column (successor regime)")

```

More than 76 per cent of transitions to democracy are from non-democracy; the rest are from situations of no authority.

## Extension 3: Transition events per year

It is possible to calculate the frequencies of transition *per year*, including the years where no transition occurred. As we see below, regime change is extremely rare: more than 97 per cent of the country years in democratic regimes, and more than 95 per cent of the country years in non-democratic regimes, do not experience transitions:

```{r table_10_1_per_year}

data <- all_gwf %>% 
  mutate(gwf_previous = ifelse(grepl("democracy",
                                     gwf_full_regimetype),
                               "Democracy",
                               ifelse(grepl("military|oligarchy|personal|party|monarchy",
                                            gwf_full_regimetype),
                                      "Non-democracy",
                                      "Other"))) %>% 
  group_by(country_name) %>%
  mutate(transition = gwf_casename != lead(gwf_casename),
         gwf_next = lead(gwf_previous),
         gwf_next = ifelse(!transition, "No transition", gwf_next)) %>%
  ungroup() %>%
  count(gwf_previous, gwf_next, sort=TRUE) %>% 
  filter(!is.na(gwf_next)) %>%
  ungroup() %>%
  mutate(total_trans = sum(n), 
         prop_trans = scales::percent_format()(n/total_trans)) 

data  %>% 
  dcast(gwf_previous ~ gwf_next, value.var="n") %>%
  select(gwf_previous, `No transition`, everything()) %>%
  knitr::kable(col.names = c("Regime",  
                             "Number of country-years with no transition", 
                             "To democracy", 
                             "To non-democracy", 
                             "To other"))

data  %>% 
  dcast(gwf_previous ~ gwf_next, value.var="prop_trans") %>%
  select(gwf_previous, `No transition`, everything()) %>%
  knitr::kable(col.names = c("Regime",  
                             "Prop. of country-years with no transition", 
                             "To democracy", 
                             "To non-democracy", 
                             "To other"))


# This is more informative by row

data %>%
  group_by(gwf_previous) %>%
  mutate(total_by_row = sum(n), prop_trans = scales::percent_format()(n/total_by_row)) %>% 
  dcast(gwf_previous ~ gwf_next, value.var="prop_trans") %>%
  select(gwf_previous, `No transition`, everything()) %>%
  knitr::kable(col.names = c("Regime",  
                             "Prop. of country-years with no transition", 
                             "To democracy", 
                             "To non-democracy", 
                             "To other"))


```

# Figure 10.1: Relationship between GDP per capita and democracy, 1800-2014

## Original figure

This figure is a scatterplot of GDP per capita vs. democracy. GDP data is the median of the per capita value from the Maddison project [@Maddison2013], the Penn World Tables [@Feenstra2013], and the [World Bank](http://data.worldbank.org/); democracy data from @Pemstein2010, extended by me [@Marquez2016]. Each point represents a country-year; each line represents a country trajectory over those years where income and democracy data exists. Highlighted points are regimes classified as monarchies by @MagaloniChuMin2013 for the 1950-2013 period, as well as other regimes in the Middle East (including regimes for the period before 1950).

```{r fig_10_1}

data <- inner_join(extended_uds,
                  economic_data) %>% 
  group_by(country_name, GWn, year, index) %>% 
  summarise(median = median(per_capita, na.rm = TRUE)) %>% 
  filter(!is.na(median))

data <- data %>% 
  left_join(magaloni) %>% 
  mutate(regime_nr = ifelse(regime_nr == "Monarchy",
                           "monarchy",
                           NA),
         region = ifelse(country_name %in%
                           c("Algeria",
                             "Bahrain",
                             "Egypt",
                             "Iraq",
                             "Israel",
                             "Jordan",
                             "Kuwait",
                             "Lebanon",
                             "Libya",
                             "Morocco",
                             "Oman",
                             "Qatar",
                             "Saudi Arabia",
                             "South Sudan",
                             "Sudan",
                             "Syria",
                             "Tunisia",
                             "Turkey (Ottoman Empire)",
                             "United Arab Emirates",
                             "Yemen (Arab Republic of Yemen)",
                             "Yemen, People's Republic of"),
                         "Middle East", 
                         ifelse(country_name == "Singapore", 
                                "Singapore", 
                                NA)), 
                        region_regime = ifelse(!is.na(regime_nr) & !is.na(region), 
                                               paste(region, regime_nr),
                                               region))


ggplot(data = data, 
       aes(x = index, 
           y = median, 
           group = GWn)) + 
  geom_path(alpha = 0.1) + 
  geom_point(aes(color = region_regime), 
             alpha = 0.5) + 
  theme_bw() + 
  scale_y_log10(label=scales::dollar) +
  geom_smooth(aes(y = median,group=0), 
              method = "lm", 
              color = "red") +
  labs(y="GDP per capita, median of various measures", 
       x="Unified democracy score",
       shape="",
       alpha="",
       color="") + 
  geom_text(data = data %>% 
              ungroup() %>% 
              filter(!is.na(region_regime),
                     (year %in% c(min(year), 
                                 max(year))) |
                       ((year %% 10) == 0)),
            aes(label = year, color = region_regime), 
            check_overlap = TRUE) + 
  geom_vline(xintercept = 0.5, linetype=2) +
  theme(legend.position = "bottom") +
  scale_color_discrete(na.value = "lightgrey") +
  guides(shape = guide_legend(ncol=2))

```

## Extension 1: Using different measures of GDP per capita and democracy

This correlation does not vary much depending on the measure of GDP per capita:

```{r fig_10_1_extension_1, fig.height = 30}

data <- inner_join(extended_uds,
                  economic_data) %>%
  filter(!is.na(per_capita))

ggplot(data = data, 
       aes(x = index, 
           y = per_capita, 
           group = GWn)) + 
  geom_path(alpha = 0.1) + 
  geom_point(alpha = 0.5) + 
  theme_bw() + 
  scale_y_log10(label=scales::dollar) +
  geom_smooth(aes(y = per_capita, group = variable),
              method = "lm",
              color = "red") +
  labs(y="GDP per capita, various measures",
       x="Unified democracy score",
       shape="",
       alpha="",
       color="") +
  geom_text(data = data %>%
              ungroup() %>%
              filter((year %in% c(min(year),
                                 max(year))) |
                       ((year %% 25) == 0)),
            aes(label = year),
            check_overlap = TRUE) +
  geom_vline(xintercept = 0.5, linetype=2) +
  theme(legend.position = "bottom") +
  scale_color_discrete(na.value = "lightgrey") +
  guides(shape = guide_legend(ncol=2)) +
  facet_wrap(~variable, ncol = 1)
```

Or democracy:

```{r fig_10_1_extension_2, fig.height = 25}

data <- full_join(vdem,
                  economic_data) %>% 
  group_by(country_name, GWn, year, 
           v2x_polyarchy, 
           v2x_api, 
           v2x_mpi, 
           v2x_libdem, 
           v2x_partipdem, 
           v2x_delibdem, 
           v2x_egaldem) %>% 
  summarise(median = median(per_capita, na.rm = TRUE)) %>% 
  filter(!is.na(median)) %>%
  melt(measure.vars = c("v2x_polyarchy",
                        "v2x_api",
                        "v2x_mpi",
                        "v2x_libdem",
                        "v2x_partipdem",
                        "v2x_delibdem",
                        "v2x_egaldem"))

ggplot(data = data, 
       aes(x = value, 
           y = median, 
           group = GWn)) + 
  geom_path(alpha = 0.1) + 
  geom_point(alpha = 0.5) + 
  theme_bw()+ 
  scale_y_log10(label=scales::dollar) +
  geom_smooth(aes(y = median, group = variable), 
              method = "lm", 
              color = "red") +
  labs(y="GDP per capita, median of various measures", 
       x="Democracy score, VDem",
       shape="",
       alpha="",
       color="") + 
  geom_text(data = data %>% 
              ungroup() %>% 
              filter((year %in% c(min(year), 
                                 max(year))) |
                       ((year %% 25) == 0)),
            aes(label = year), 
            check_overlap = TRUE) + 
  geom_vline(xintercept = 0.5, linetype=2) +
  theme(legend.position = "bottom") +
  scale_color_discrete(na.value = "lightgrey") +
  guides(shape = guide_legend(ncol=2)) +
  facet_wrap(~variable, ncol = 1)
```

## Extension 2: Country by country correlation

But the correlation does vary from country to country, and depends a bit on the measure of GDP used:

```{r fig_10_1_extension_3, fig.height=30}

library(broom)
library(purrr)
library(tidyr)

data <- left_join(extended_uds,
                  economic_data) %>% 
  group_by(primary_source, variable, country_name) %>% 
  filter(!is.na(per_capita), n() > 5) %>% 
  summarise(index = list(index), per_capita = list(per_capita)) %>%
  mutate(cor_test = map2(index, per_capita, ~ tidy(cor.test(.x, log(.y))))) %>% 
  unnest(cor_test) %>%
  ungroup()

ggplot(data = data,
       aes(x = forcats::fct_reorder(country_name, estimate, na.rm =TRUE),
           y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high, 
                      color = primary_source)) +
  labs(x = "",
       y = "Estimated correlation",
       color = "Primary source of GDP data") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol=2)) +
  geom_hline(yintercept = 0, color = "red") +
  coord_flip()


```

We can also do this as follows:

```{r fig_10_1_extension_3_1, fig.height=30}

data <- left_join(extended_uds,
                  economic_data) %>% 
  group_by(primary_source, variable, country_name) %>% 
  filter(!is.na(per_capita), n() > 5) %>% 
  summarise(index = list(index), per_capita = list(per_capita)) %>%
  mutate(cor_test = map2(index, per_capita, ~ tidy(cor.test(.x, log(.y))))) %>% 
  unnest(cor_test) %>% 
  group_by(country_name) %>%
  summarise(avg_estimate = mean(estimate),
            se_estimate = sqrt(sum(((estimate - conf.high)/1.96)^2))/n(),
            conf.high = avg_estimate + 1.96*se_estimate,
            conf.low = avg_estimate - 1.96*se_estimate)

ggplot(data = data,
       aes(x = forcats::fct_reorder(country_name, avg_estimate, na.rm =TRUE),
           y = avg_estimate)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high)) +
  labs(x = "",
       y = "Estimated correlation",
       color = "Primary source of GDP data") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol=2)) +
  geom_hline(yintercept = 0, color = "red") +
  coord_flip()

```

# Figure 10.2: Year by year correlation between economic development and democracy

Raw (uncontrolled) correlation between economic development and democracy, per year. Shaded areas represent 95% confidence intervals. GDP data from the Maddison Project [@Maddison2013], the Penn World Tables [@Feenstra2013], and the [World Bank](http://data.worldbank.org/); democracy data from @Pemstein2010, extended by the me [@Marquez2016].

```{r fig_10_2}

data <- left_join(extended_uds,
                  economic_data) %>% 
  group_by(primary_source, variable, year) %>% 
  filter(!is.na(per_capita), n() > 4) %>% 
  summarise(index = list(index), per_capita = list(per_capita)) %>%
  mutate(cor_test = map2(index, per_capita, ~ tidy(cor.test(.x, log(.y))))) %>% 
  unnest(cor_test) %>% 
  group_by(year) %>% 
  summarise(median = median(estimate, na.rm=TRUE),
         min_conf_low = min(conf.low, na.rm=TRUE),
         max_conf_high = max(conf.high, na.rm=TRUE))

ggplot(data = data, 
       aes(x = year, 
           y = median, 
           ymin = min_conf_low, 
           ymax = max_conf_high)) + 
  geom_path() + 
  theme_bw()+ 
  geom_ribbon(aes(ymax = max_conf_high, 
                  ymin = min_conf_low),
              alpha = 0.3, fill = "lightgrey") + 
  geom_smooth(aes(y=median), color = "black") +
  labs(y="Raw correlation between
       GDP per capita (various measures) and
       extended Unified Democracy Score, per year") + 
  geom_hline(yintercept = 0, linetype=2) +
  coord_cartesian(xlim=c(1850,2014))

```

# Figure 10.3: Inequality in Egypt, 1964-2009

Inequality data is from the Standardized Worldwide Income Inequality Database [@Solt2016; @Solt2009].

```{r fig_10_3}
count_sequence_breaks <- function(seq, seq_step = 1) {
  first_diff <- c(seq_step, diff(seq)) - seq_step
  periods <- cumsum(abs(first_diff))
  periods
}

data <- swiid_summary_5 %>% 
  filter(country_name  %in% c("Egypt","Tunisia")) 

world_ineq <- swiid_summary_5 %>% 
  group_by(year,variable) %>% 
  summarise(mean_value = mean(mean_value,na.rm=TRUE), 
            se = sqrt(sum(se^2))/n())

middle_east_ineq <- swiid_summary_5 %>% 
  filter(region %in% c("Northern Africa","Western Asia"), 
         !(country_name %in% c("Georgia","Azerbaijan","Cyprus"))) %>% 
  group_by(year,variable) %>% 
  summarise(mean_value = mean(mean_value,na.rm=TRUE), 
            se = sqrt(sum(se^2))/n())

world_ineq$country_name <- "World avg."

middle_east_ineq$country_name <- "Middle East avg."

data <- bind_rows(data, world_ineq, middle_east_ineq) %>%
  filter(variable == "gini_net") %>%
  group_by(country_name) %>%
  arrange(year) %>%
  mutate(period = paste(count_sequence_breaks(year), country_name)) %>%
  ungroup()

ggplot(data = data %>% filter(country_name != "Tunisia"), 
       aes(x=year, 
           y = mean_value, 
           color = country_name, 
           fill = country_name, 
           ymin = mean_value - 1.96*se, 
           ymax = mean_value + 1.96*se, 
           group = period)) + 
  geom_path() + 
  geom_point() +
  geom_ribbon(alpha=0.2) +
  labs(fill = "",
       color ="", 
       y = "Net gini index\n(after taxes and transfers,
       higher values indicate more inequality)") +
  theme_bw() +
  theme(legend.position ="bottom") +
  geom_vline(xintercept = 2011, color = "red") +
  geom_text(x = 2011,
            y = 50,
            label = "2011 revolution", 
            angle = 90,
            color = "black")

```

# Figure 10.4: Inequality in Tunisia, 1965-2010

```{r fig_10_4}

ggplot(data = data %>% filter(country_name != "Egypt"), 
       aes(x=year, 
           y = mean_value, 
           color = forcats::fct_relevel(country_name, "Tunisia"), 
           fill = forcats::fct_relevel(country_name, "Tunisia"), 
           ymin = mean_value - 1.96*se, 
           ymax = mean_value + 1.96*se, 
           group = period)) + 
  geom_point() +
  geom_path() + 
  geom_ribbon(alpha=0.2) +
  labs(fill = "",
       color ="", 
       y = "Net gini index\n(after taxes and transfers,
       higher values indicate more inequality)") +
  theme_bw() +
  theme(legend.position ="bottom") +
  geom_vline(xintercept = 2011, color = "red") +
  geom_text(x = 2011,
            y = 50,
            label = "2011 revolution", 
            angle = 90,
            color = "black")

```

# Figure 10.5: Inequality in Venezuela

```{r fig_10_5}

data <- swiid_summary_5 %>% 
  filter(country_name  %in% c("Venezuela"))

latinamerica_ineq <- swiid_summary_5 %>% 
  filter(region %in% c("Southern America","Central America")) %>% 
  group_by(year, variable) %>% 
  summarise(mean_value = mean(mean_value, na.rm=TRUE), 
            se = sqrt(sum(se^2))/n())

latinamerica_ineq$country_name <- "Latin America avg."

data <- bind_rows(data, world_ineq, latinamerica_ineq) %>% 
  filter(variable == "gini_net") %>%
  group_by(country_name) %>%
  arrange(year) %>%
  mutate(period = paste(count_sequence_breaks(year), country_name)) %>%
  ungroup()

ggplot(data = data %>% filter(country_name != "Egypt"), 
       aes(x=year, 
           y = mean_value, 
           color = forcats::fct_relevel(country_name, "Venezuela"), 
           fill = forcats::fct_relevel(country_name, "Venezuela"), 
           ymin = mean_value - 1.96*se, 
           ymax = mean_value + 1.96*se, 
           group = period)) + 
  geom_path() + 
  geom_point() +
  geom_ribbon(alpha=0.2) +
  labs(fill = "",
       color ="", 
       y = "Net gini index\n(after taxes and transfers,
       higher values indicate more inequality)") +
  theme_bw() +
  theme(legend.position ="bottom") +
  geom_vline(xintercept = 1997, color = "red") +
  geom_text(x = 1997,
            y = 50,
            label = "Chavez is elected", 
            angle = 90,
            color = "black")
```

# Figure 10.6: Inequality and Political Regimes

## Original figure

The inequality measure is the net gini index of income inequality (net means after taxes and transfers); it goes from 0 to 100, where higher numbers mean more inequality. Data on regimes is from @Kailitz2013. Inequality data from Solt [@Solt2009; @Solt2016]. The democracy data is from @Pemstein2010, extended by me [@Marquez2016]. The original figure used the package `ggalt`, which is no longer compatible with `ggplot2`; nothing changes if we use `ggplot2::geom_density_2d` for the estimation.

```{r fig_10_6}

data <- inner_join(swiid_summary_5,
                   extended_uds %>% 
                     select(country_name, GWn, year, index) %>% 
                     rename(uds = index)) %>%
  inner_join(kailitz_yearly %>%
               select(-cown:-in_system) %>%
               mutate(democracy = grepl("Democracy",combined_regime)) %>%
               melt(measure.vars = c("personal",
                                     "communist",
                                     "military",
                                     "party",
                                     "monarchy",
                                     "electoral",
                                     "failure",
                                     "transition",
                                     "democracy")) %>%
               filter(value) %>%
               rename(regime = variable)) %>%
  filter(!is.na(regime),
         variable %in% c("gini_net")) %>%
  mutate(name = paste(country_name, year))


ggplot(data = data, 
       aes(text = name,
           x = uds,
           y = mean_value,
           group = regime)) +
  geom_point(alpha = 0.2) +
  geom_density_2d() +
  theme_bw() +
  labs(y = "Gini index",
       x = "Unified democracy score (democracies = >0.5)",
       alpha = "Concentration of regimes",
       fill = "") +
  theme(legend.position = "bottom") + 
  geom_vline(xintercept = 0.5, color = "red") +
  facet_wrap(~regime) +
  coord_cartesian(xlim=c(0,1))

```

### Interactive version

Here's an interactive version (requires the `plotly` package):

```{r fig_10_6_interactive}
library(plotly)

p <- ggplot(data = data %>%
              rename(gini_net = mean_value), 
       aes(text = name,
           x = uds,
           y = gini_net,
           group = regime)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  labs(y = "Gini index",
       x = "Unified democracy score (democracies = >0.5)",
       alpha = "Concentration of regimes",
       fill = "") +
  theme(legend.position = "bottom") + 
  geom_vline(xintercept = 0.5, color = "red") +
  facet_wrap(~regime) +
  coord_cartesian(xlim=c(0,1))

ggplotly(p)

```


## Extension 1: Other measures of regime type

We can use the measure of regime type from @MagaloniChuMin2013 instead:

```{r fig_10_6_ext_1}

data <- inner_join(swiid_summary_5,
                   extended_uds %>% 
                     select(country_name,
                            GWn,
                            year,
                            index) %>% 
                     rename(uds = index)) %>%
  inner_join(magaloni %>% 
                     select(country_name,
                            GWn,
                            year,
                            regime_nr)) %>%
  left_join(kailitz_yearly %>%
              select(country_name, 
                     GWn, 
                     year, 
                     communist)) %>%
  mutate(regime = ifelse(communist, 
                         paste("Communist ", regime_nr), 
                         regime_nr)) %>%
  filter(!is.na(regime),
         variable %in% c("gini_net"))


ggplot(data = data, 
       aes(x = uds,
           y = mean_value,
           group = regime)) +
  geom_point(alpha = 0.2) +
  geom_density_2d() +
  theme_bw() +
  labs(y = "Gini index",
       x = "Unified democracy score (democracies = >0.5)",
       alpha = "Concentration of regimes",
       fill = "") +
  theme(legend.position = "bottom") + 
  geom_vline(xintercept = 0.5, color = "red") +
  facet_wrap(~regime) +
  coord_cartesian(xlim=c(0,1))


```

Or from @WahmanTeorellHadenius2013:

```{r fig_10_6_ext_2}

data <- inner_join(swiid_summary_5,
                   extended_uds %>% 
                     select(country_name,
                            GWn,
                            year,
                            index) %>% 
                     rename(uds = index)) %>%
  inner_join(wahman_teorell %>% 
                     select(country_name,
                            GWn,
                            year,
                            regime1ny)) %>%
  filter(!is.na(regime1ny)) %>%
  left_join(kailitz_yearly %>%
              select(country_name, 
                     GWn, 
                     year, 
                     communist)) %>%
  mutate(regime = ifelse(communist, 
                         paste("Communist ", regime1ny), 
                         as.character(regime1ny))) %>%
  filter(!is.na(regime),
         variable %in% c("gini_net"))

ggplot(data = data, 
       aes(x = uds,
           y = mean_value,
           group = regime)) +
  geom_point(alpha = 0.2) +
  geom_density_2d() +
  theme_bw() +
  labs(y = "Gini index",
       x = "Unified democracy score (democracies = >0.5)",
       alpha = "Concentration of regimes",
       fill = "") +
  theme(legend.position = "bottom") + 
  geom_vline(xintercept = 0.5, color = "red") +
  facet_wrap(~regime) +
  coord_cartesian(xlim=c(0,1))


```

Or from @GeddesWrightFrantz2014:

```{r fig_10_6_ext_3, fig.height = 8}

data <- inner_join(swiid_summary_5,
                   extended_uds %>% 
                     select(country_name,
                            GWn,
                            year,
                            index) %>% 
                     rename(uds = index)) %>%
  inner_join(all_gwf %>% 
                     select(country_name,
                            GWn,
                            year,
                            gwf_full_regimetype)) %>%
  filter(!is.na(gwf_full_regimetype)) %>%
  left_join(kailitz_yearly %>%
              select(country_name, 
                     GWn, 
                     year, 
                     communist)) %>%
  mutate(regime = ifelse(communist, 
                         paste("Communist ", gwf_full_regimetype), 
                         as.character(gwf_full_regimetype))) %>%
  filter(!is.na(regime),
         variable %in% c("gini_net"))

ggplot(data = data, 
       aes(x = uds,
           y = mean_value,
           group = regime)) +
  geom_point(alpha = 0.2) +
  geom_density_2d() +
  theme_bw() +
  labs(y = "Gini index",
       x = "Unified democracy score (democracies = >0.5)",
       alpha = "Concentration of regimes",
       fill = "") +
  theme(legend.position = "bottom") + 
  geom_vline(xintercept = 0.5, color = "red") +
  facet_wrap(~regime) +
  coord_cartesian(xlim=c(0,1))


```

## Extension 2: Country-level correlations between inequality and democracy level

The correlation between inequality and democracy has varied accross countries.

```{r ineq_countries, fig.height=25}

data <- left_join(extended_uds,
                  swiid_summary_5 %>% 
                    select(country_name, 
                           GWn, 
                           year, 
                           variable, 
                           mean_value)) %>%
  filter(variable == "gini_net") %>%
  group_by(variable, country_name) %>% 
  filter(!is.na(mean_value), 
         n() > 5) %>% 
  summarise(index = list(index), mean_value = list(mean_value)) %>%
  mutate(cor_test = map2(index, mean_value, ~ tidy(cor.test(.x, .y)))) %>% 
  unnest(cor_test)

ggplot(data = data,
       aes(x = forcats::fct_reorder(country_name, estimate, na.rm =TRUE),
           y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, 
                      ymax = conf.high)) +
  labs(x = "",
       y = "Estimated correlation",
       color = "Measure") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(ncol=2)) +
  geom_hline(yintercept = 0, color = "red") +
  coord_flip()

```

# References
