---
title: "Chapter 3: The Struggle over Social Control: Totalitarian and Authoritarian Rule"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
bibliography: bibfile.bib
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Chapter 3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette shows how to replicate and extend the charts in chapter 3 of my book [*Non-democratic Politics: Authoritarianism, Dictatorships, and Democratization*](https://he.palgrave.com/page/detail/non-democratic-politics-xavier-marquez/?sf1=barcode&st1=9781137486301) (Palgrave Macmillan, 2016). It assumes that you have downloaded the replication package as follows:

```{r, echo = TRUE, eval = FALSE}
if(!require(devtools)) {
  install.packages("devtools")
}

devtools::install_github('xmarquez/AuthoritarianismBook')

```

It also assumes you have the `dplyr`, `ggplot2`, `scales`, `forcats`, and `knitr` packages installed:

```{r, echo = TRUE, eval = FALSE}
if(!require(dplyr)) {
  install.packages("dplyr")
}

if(!require(ggplot2)) {
  install.packages("ggplot2")
}

if(!require(scales)) {
  install.packages("forcats")
}

if(!require(forcats)) {
  install.packages("scales")
}

if(!require(knitr)) {
  install.packages("knitr")
}
```

```{r setup_chapter_3, include=FALSE}
knitr::opts_chunk$set(fig.height = 7, 
                      fig.width = 7)
```

## Figure 3.1: Restrictions on Party Competition in Non-democratic Regimes

### Original figure

This figure shows restrictions on party competition in  non-democratic regimes in the post-WWII era. Excludes countries with fewer than 500,000 people). Data from [@Svolik2012]. Svolik counts "united fronts" as single parties when the parties involved are not genuinely independent, as in China and the GDR, where ruling communist parties supposedly rule (or ruled, in the case of the GDR) in a "united front" with other parties.

```{r fig_3_1}
library(AuthoritarianismBook)
library(dplyr)
library(ggplot2)

data <- svolik_institutions

ggplot(data = data, aes(x = year, fill = party)) + 
  geom_bar(position = "fill", width = 1) +
  theme_bw() +
  labs(fill = "Party restrictions",
       x = "Year", 
       y = "Proportion of non-democratic regimes") + 
  guides(color = guide_legend(title.position="top")) +
  theme(legend.position = "bottom") +
  geom_vline(xintercept = 1989, linetype = 2) +
  scale_y_continuous(label=scales::percent) + 
  annotate("text", 
           x = 1989,
           y = 0.5,
           label = "End of the Cold War", angle=90)

```

### Extension: Disaggregating by country

We can also see the data per country:

```{r per_country, fig.height = 20}

ggplot(data = data, 
       aes(x = forcats::fct_rev(reorder(as.factor(country_name), 
                                        year, 
                                        FUN = min)), 
           y = year))  +
  geom_tile(data = data, 
             aes(fill = party)) +
  theme_bw() +
  coord_flip() +
  labs(x = "", fill = "Party restrictions") + 
  theme(legend.position = "top") +
  guides(fill = guide_legend(ncol = 1, title.position="top")) +
  geom_hline(yintercept = 1989, linetype = 2) +
  annotate("text", 
           y = 1989,
           x = length(unique(data$country_name))/2,
           label = "End of the Cold War", angle=90)

```


## Other Extensions: Measuring Totalitarianism

Though the chapter does not use a measure of totalitarianism, it is possible to construct one using the civil society measure (`v2xcs_ccsi`) from V-Dem [@vdem2016dataset; @Benrhardt2015] discussed in the vignette for [chapter 1](Chapter_1.html):

```{r totalitarian_regimes_prep}

civil_society <- vdem %>% 
  select(country_name, GWn, year, 
         in_system, v2xcs_ccsi, 
         v2xcs_ccsi_codelow, v2xcs_ccsi_codehigh) %>%
  mutate(civil_society = qnorm(1 - v2xcs_ccsi),
         civil_society_high = qnorm(1 - v2xcs_ccsi_codelow), 
         civil_society_low = qnorm(1 - v2xcs_ccsi_codehigh))

```

Countries that are at least two standard deviations away from the mean in this measure can be considered totalitarian, and countries that are 1.5 standard deviations away may be close to totalitarian. Here are they:

```{r totalitarian_regimes, fig.height = 9}

data <- civil_society %>%
  filter(in_system, civil_society >= 1.5) %>%
  inner_join(kailitz_yearly) %>%
  mutate(totalitarian = forcats::fct_rev(factor((civil_society >= 2),
                               labels = c("Near-totalitarian","Totalitarian"))),
         communist = ifelse(communist, "communist", "non-communist"))

ggplot(data = data, 
       aes(x = forcats::fct_rev(reorder(as.factor(country_name), 
                                        -civil_society, 
                                        FUN = min)), 
           y = year))  +
  geom_tile(aes(fill = civil_society)) +
  theme_bw() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 1.75) +
  coord_flip() +
  labs(x = "", 
       fill = "Degree of control over civil society") + 
  theme(legend.position = "top") +
  guides(color = guide_legend(ncol = 1, title.position="top"),
         size = guide_legend(ncol = 1, title.position="top")) +
  facet_grid(communist ~ totalitarian, scales = "free_y")

```

This measure may be averaged with Fariss' measure of repression [@Fariss2014] for a different look:

```{r totalitarian_regimes_2, fig.height = 9}

data <- inner_join(civil_society, fariss %>% 
                    select(country_name, year, GWn, latentmean)) %>%
  inner_join(kailitz_yearly %>% 
                    select(country_name, year, GWn, communist)) %>%
  mutate(totalitarian_score = (civil_society - latentmean)/2,
         totalitarian = factor(totalitarian_score >= 2, 
                               labels = c("Near-totalitarian", "Totalitarian")),
         communist = ifelse(communist, "communist", "non-communist"))

ggplot(data = data %>% filter(totalitarian_score >= 1.5),
       aes(x = forcats::fct_rev(reorder(as.factor(country_name),
                                        -totalitarian_score,
                                        FUN = min)),
           y = year))  +
  geom_tile(aes(fill = totalitarian_score)) +
  theme_bw() +
  scale_fill_gradient2(low = "blue", high = "red", midpoint = 1.75) +
  coord_flip() +
  labs(x = "",
       fill = "Degree of control over civil society\naveraged with degree of repression") +
  theme(legend.position = "top") +
  guides(color = guide_legend(ncol = 1, title.position="top"),
         size = guide_legend(ncol = 1, title.position="top")) +
  facet_grid(communist ~ totalitarian, scales = "free_y")

```

# References
